---
title: "Lab6 - GeoTIFF, kriging"
author: "Robert Saldivar"
date: "February 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)
```

## Grand Canyon GeoTIFF

```{r}

gc_dem <- raster(here("data", "gc_dem.tif"))


# Look at it using plot():
plot(gc_dem)

# check the CRS
gc_dem@crs

# Check the extent (bounding box):

gc_dem@extent

# Creating a wgs84 w/latlong:
wgs84 <- "+proj=longlat + datum=WGS84 +ellps=WGS84 +no_defs"

gc_reproj <- projectRaster(gc_dem, crs = wgs84, method = "bilinear") 

gc_reproj@extent

```

### Crop raster to a smaller area (bounding box):
```{r}
bounds <- as(extent(-112.4, 112.0, 36.1, 36.3), 'SpatialPolygons')
```

### Make the CRS of our bounding box the same as for gc_reproj:
```{r}
crs(bounds) <- crs(gc_reproj)
```

### Now let's crop our original spatial data:
```{r}
gc_crop <- crop(gc_reproj, bounds)

plot(gc_crop)
```

### Resample using the 'aggregate()' function:
```{r}
gc_agg <- raster::aggregate(gc_crop, fact = 10)

plot(gc_agg)
```

### Now let's get safe and sound in ggplot:

First, convert data to a data frame:
```{r}
gc_df <- as.data.frame(gc_agg, xy=TRUE)

ggplot(data = gc_df, aes(x = x, y = y)) +
  geom_raster(aes(fill = layer)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c(
    "purple",
    "magenta",
    "orange",
    "yellow",
    "white"
  ))
```

### HOw can I just select cells that match given criteria
```{r}
gc_hab <- gc_crop

# Set any cells outside of [1000 - 1500] to NA

gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

# Plot with base plot:
plot(gc_hab)

```

Now let's make this interactive with tmap:
```{r}
tmap_mode("view")

tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")

```

### Kriging rain in Kansas

Read in the KS counties shapefile data
```{r}
ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp"))

# Use base plot:
plot(ks_counties)

#check CRS:
st_crs(ks_counties) #NA, let's set one!

# Set to EPSG 4326:
st_crs(ks_counties) <- 4326

plot(ks_counties)
```

Practice in ggplot:
```{r}
ggplot(data = ks_counties) +
  geom_sf()
```

Now, read in the rainfall data:
```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>%
  clean_names()
```

Update the ks_rain data to be recognized as spatial points:
```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326)
```

```{r}
ggplot() + 
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf, 
          aes(color = amt, size = amt))
```

## Kriging to predict rainfall:
```{r}
ks_sp <- as_Spatial(ks_sf)
class(ks_sp)
```

### Make a spatial pixels grid taht we'll make predictions over:

```{r}
lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6, 102, length.out = 200)

# Now we're going to make this into a spatial grid:
grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

plot(grid_sp)
```

### Make a variogram
```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

# My estimates for variogram parameters:
# nugget = 0.2
# sill = 0.8
# range = 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))

plot(ks_vgm, ks_vgm_fit)
```

### Now, krige!
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)
```

```{r}
spplot(ks_krige, "var1.pred")
```

 ### Make a data frame of the kriged predictions
```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>%
  rename(longitude = coords.x1,
         latitude = coords.x2)

# Converting this df to sf object:
rain_sf <- st_as_sf(ks_df, coords = c("longitude", "latitude"), crs = 4326)

ggplot(rain_sf) +
  geom_sf(aes(color = var1.pred))
```
 
 
 ### Crop to the actual KS outline:
```{r}
ks <- read_sf(dsn = here("data", "states"),
              layer = "cb_2017_us_state_20m") %>%
  dplyr::select(NAME) %>%
  filter(NAME == "Kansas") %>%
  st_transform(crs = 4326)

plot(ks)
```
 
```{r}
#Find the intersection of the two:
rain_sf_ks <- st_intersection(rain_sf, ks) 

#Plot!

ggplot(data = rain_sf_ks) +
  geom_sf(aes(color = var1.pred))

```
 
 
